---
title: "Project 3 - Example Main Script"
author: "Group5: Peifeng Hong, Leo Lam, Shiyu Liu, Qianli Zhu, Sitong Chen"
date: "20180322"
output: html_document
---
Package needed
```{r "package import", message = FALSE, warnings = FALSE}
#packages <- c("e1071", "dplyr","gbm","randomForest","EBImage","xgboost","data.table")


#packages.needed=setdiff(packages,intersect(installed.packages()[,1], packages))

#if(length(packages.needed)>0){install.packages(packages.needed, dependencies = TRUE)}

library("e1071")
library("dplyr")
library("gbm")
library("randomForest")
library("EBImage")
library("xgboost")
library("data.table")
```


### Step 0: specify directories.

Set the working directory to the image folder.

```{r wkdir, eval=FALSE}
setwd("~/Documents/GitHub/Spring2018-Project3-Group5")
```
Provide directories for raw images. Training set and test set should be in different subfolders. 
```{r}
experiment_dir <- "~/Documents/GitHub/Spring2018-Project3-Group5/data/" # This will be modified for different data sets.
img_train_dir <- paste(experiment_dir, "train/images/", sep="")
img_test_dir <- paste(experiment_dir, "test/images/", sep="")
```

### Step 1: set up controls for evaluation experiments.

In this chunk, ,we have a set of controls for the evaluation experiments. 

+ (T/F) cross-validation on the training set
+ (number) K, the number of CV folds
+ (T/F) process features for training set
+ (T/F) run evaluation on an independent test set
+ (T/F) process features for test set

```{r exp_setup}
K <- 5  # number of CV folds
run.feature.train=F # process features for training pictures
run.feature.test=T # process features for testing pictures
run.train = F # if true, train model on training data, else use saved model
run.test=T # run evaluation on an independent test set
run.train.baseline=F
```


### Step 2: import training images class labels.

```{r train_label}
labels<-read.csv("~/Documents/GitHub/Spring2018-Project3-Group5/output/label_train.csv")
label_train<-labels$label
```



### Step 3: Construct visual features
```{r feature}
source("~/Documents/GitHub/Spring2018-Project3-Group5/lib/feature.R")

  
if(run.feature.train){

# get SIFT feature
  sift_feature_train <- read.csv("~/Documents/GitHub/Spring2018-Project3-Group5/output/SIFT_feature.csv",header = F)
  sift_feature_train <- sift_feature_train[,-1]
  sift_feature_train$y <- label_train
  saveRDS(sift_feature_train, file="~/Documents/GitHub/Spring2018-Project3-Group5/output/feature_SIFT_train.RData")
  
# get RGB feature
  tm_feature <- system.time(rgb_feature_train <- feature(img_train_dir, 
        "train", 
        data_name="rgb", 
        export=TRUE))[3]
} else {
#load SIFT feature
  sift_feature_train<-readRDS("~/Documents/GitHub/Spring2018-Project3-Group5/output/feature_SIFT_train.RData")
#load RGB feature
  rgb_feature_train<-readRDS('~/Documents/GitHub/Spring2018-Project3-Group5/output/feature_rgb_train.RData')

}


if(run.feature.test){

# get SIFT feature
  sift_feature_test <- read.csv("~/Documents/GitHub/Spring2018-Project3-Group5/output/SIFT_test.csv",header = F)
  sift_feature_test <- sift_feature_test[,-1]
  saveRDS(sift_feature_test, file="~/Documents/GitHub/Spring2018-Project3-Group5/output/feature_SIFT_test.RData")

# get RGB feature
  tm_feature_test <- system.time(rgb_feature_test <- feature(img_test_dir, 
        "test", 
        data_name="rgb", 
        export=TRUE))[3]
  saveRDS(rgb_feature_test, file="~/Documents/GitHub/Spring2018-Project3-Group5/output/feature_rgb_test.RData")
  
}
```

### Step 4: Train a baseline model (GBM+SIFT) with training images

Call the train model and test model from library. 
```{r loadlib}
source("~/Documents/GitHub/Spring2018-Project3-Group5/lib/train_gbm.R")
source("~/Documents/GitHub/Spring2018-Project3-Group5/lib/test_gbm.R")
#source("~/Documents/GitHub/Spring2018-Project3-Group5/lib/cross_validation.R")
```
```{r}
if(run.train.baseline){ 
# process data
  training_y_SIFT <- label_train
  training_x_SIFT<- sift_feature_train[ ,!(colnames(sift_feature_train) %in% c("y"))]
  
# train model
  tm_train_base <- system.time(base_model <-  
              train(training_x_SIFT,training_y_SIFT))[3]
  saveRDS(base_model,file='~/Documents/GitHub/Spring2018-Project3-Group5/output/base_model.RData')
}else{
  base_model <- readRDS('~/Documents/GitHub/Spring2018-Project3-Group5/output/base_model.RData')
}
base_model <- readRDS('~/Documents/GitHub/Spring2018-Project3-Group5/output/base_model.RData')
```


### Step 5: Train advanced model (XGBoost+RGB) with training images
```{r loadslib}
source("~/Documents/GitHub/Spring2018-Project3-Group5/lib/train_xgboost.r")
source("~/Documents/GitHub/Spring2018-Project3-Group5/lib/test_xgboost.r")
```
```{r}
if(run.train){
# process data
  training_y_RGB<- label_train-1
  training_x_RGB<- rgb_feature_train[,!(colnames(rgb_feature_train) %in% c("y"))]
# train model
  best_para <- xgboost_para(rgb_feature_train,training_y_RGB,K)[[2]]
  tm_train <- system.time(best_model <- xgboost_train(training_x_RGB, training_y_RGB, best_para))[3]
  saveRDS(best_model,file='~/Documents/GitHub/Spring2018-Project3-Group5/output/best_model.RData')
  best_model <- readRDS('~/Documents/GitHub/Spring2018-Project3-Group5/output/best_model.RData')
}
best_model <- readRDS('~/Documents/GitHub/Spring2018-Project3-Group5/output/best_model.RData')
```

### Step 6: Make prediction 
Feed the final training model with the completely holdout testing data. 
```{r test}
tm_test_base=NA
tm_test=NA
if(run.test){
  #process data
  testing_x_SIFT<- sift_feature_test
  #test model
  tm_test_base <- system.time(base_test_model <- test(base_model,testing_x_SIFT))[3]
  #saveRDS(base_test_model,'~/Documents/GitHub/Spring2018-Project3-Group5/output/labelsbase.Rdata')
  write.csv(file='~/Documents/GitHub/Spring2018-Project3-Group5/output/best_test_model_baseline.csv', x=base_test_model)
}
if(run.test){
  testing_x_rgb<- rgb_feature_test
  #best_test_model is the result
  tm_test <- system.time(best_test_model <- xgboost_test_result(best_model, rgb_feature_test)+1)[3]
  #saveRDS(best_test_model,file='~/Documents/GitHub/Spring2018-Project3-Group5/output/best_test_model.RData')
  write.csv(file='~/Documents/GitHub/Spring2018-Project3-Group5/output/best_test_model_advanved.csv', x=best_test_model)
}
```

### Summarize Running Time
Prediction performance matters, so does the running times for constructing features and for training the model, especially when the computation resource is limited. 
```{r running_time}
cat("Time for constructing test features=", tm_feature_test, "s \n")
#cat("Time for constructing testing features=", tm_feature_test, "s \n")
#cat("Time for training baseline model=", tm_train_base, "s \n")
cat("Time for training advanced model=", tm_train, "s \n")
cat("Time for baseline model to make prediction=", tm_test_base, "s \n")
cat("Time for advanced model to make prediction=", tm_test, "s \n")
#accuracy_test <- mean(best_test_model != label_train)
#cat("Accuracy rate for advanced model=",accuracy_test,"s \n")
```
