---
title: "R Notebook"
output: html_notebook
---

#############################################################
### train Random Forest model for training/testing images ###
#############################################################

### Authors: Peifeng Hong
### Project 3
### ADS Spring 2018

### Step1: Train and Test Function

```{r}

train<- function(dat, label,N){
  
  library(randomForest)
  
  t = proc.time()
  df <- as.data.frame(cbind(dat, label))
  colnames(df)[ncol(df)]<-"y"
  fit <- randomForest(as.factor(y)~.,data = df, importance = TRUE,ntree = N)
  
  train_time = (proc.time() - t)[3]
  cat("Elapsed time for Training Random Forest with 500 trees is ", train_time, " seconds \n")
  
  return(fit)
}


test<- function(model, dat){
  
  pred <-predict(model,newdata=dat)
 
  return (pred)
}

```


### Step2: Chooce the best parameter

```{r}

#lbp data
training <- read.csv("../output/lbp_training.csv")[,-1]
testing <- read.csv("../output/lbp_testing.csv")[,-1]

training_y <- training[ , "y"]
training_x <- training[ ,!(colnames(training) %in% c("y"))]
testing_y <- testing[ ,"y"]
testing_x <- testing[ , !(colnames(testing) %in% c("y"))]

#parameter choosing
error_testing<-c()
N<-c(300,400,500,600,700,800,900)

for(i in 1:7){

  fit <- train(training_x,training_y,N[i])

  pred_testing <- test(fit, testing_x)

  error_testing[i] <- mean(pred_testing != as.factor(testing_y))
  cat("Accurate rate for Training Random Forest with " ,N[i]," trees is ", 1-error_testing[i],'\n' )

}



```
Elapsed time for Training Random Forest with  300  trees is  28.39  seconds 
Accurate rate for Training Random Forest with  300  trees is  0.8344444
Elapsed time for Training Random Forest with  400  trees is  36.77  seconds 
Accurate rate for Training Random Forest with  400  trees is  0.8433333
Elapsed time for Training Random Forest with  500  trees is  45.82  seconds 
Accurate rate for Training Random Forest with  500  trees is  0.8455556
Elapsed time for Training Random Forest with  600  trees is  57.04  seconds 
Accurate rate for Training Random Forest with  600  trees is  0.8444444
Elapsed time for Training Random Forest with  700  trees is  67.42  seconds 
Accurate rate for Training Random Forest with  700  trees is  0.85
Elapsed time for Training Random Forest with  800  trees is  74.27  seconds 
Accurate rate for Training Random Forest with  800  trees is  0.85
Elapsed time for Training Random Forest with  900  trees is  83.63  seconds 
Accurate rate for Training Random Forest with  900  trees is  0.8411111

Best parameter for lbp feature + RandomForest should be 700.

```{r}
#hog data
training <- read.csv("../output/hog_training.csv")
testing <- read.csv("../output/hog_testing.csv")

training_y <- training[ , "y"]
training_x <- training[ ,!(colnames(training) %in% c("y"))]
testing_y <- testing[ ,"y"]
testing_x <- testing[ , !(colnames(testing) %in% c("y"))]

#parameter choosing
error_testing<-c()
N<-c(300,400,500,600,700,800,900)

for(i in 1:7){

  fit <- train(training_x,training_y,N[i])

  pred_testing <- test(fit, testing_x)

  error_testing[i] <- mean(pred_testing != as.factor(testing_y))
  cat("Accurate rate for Training Random Forest with " ,N[i]," trees is ", 1-error_testing[i],'\n' )

}
```
Elapsed time for Training Random Forest with  300  trees is  57.67  seconds 
Accurate rate for Training Random Forest with  300  trees is  0.7988889 
Elapsed time for Training Random Forest with  400  trees is  75.55  seconds 
Accurate rate for Training Random Forest with  400  trees is  0.7944444 
Elapsed time for Training Random Forest with  500  trees is  94.76  seconds 
Accurate rate for Training Random Forest with  500  trees is  0.8111111 
Elapsed time for Training Random Forest with  600  trees is  115.47  seconds 
Accurate rate for Training Random Forest with  600  trees is  0.8088889 
Elapsed time for Training Random Forest with  700  trees is  130.89  seconds 
Accurate rate for Training Random Forest with  700  trees is  0.8077778 
Elapsed time for Training Random Forest with  800  trees is  150.78  seconds 
Accurate rate for Training Random Forest with  800  trees is  0.8088889 
Elapsed time for Training Random Forest with  900  trees is  169.95  seconds 
Accurate rate for Training Random Forest with  900  trees is  0.8022222 

Best parameter for HoG feature + RandomForest is 500.


```{r}
#hsv data
training <- read.csv("../output/hsv_training.csv")[,-1]
testing <- read.csv("../output/hsv_testing.csv")[,-1]

training_y <- training[ , "y"]
training_x <- training[ ,!(colnames(training) %in% c("y"))]
testing_y <- testing[ ,"y"]
testing_x <- testing[ , !(colnames(testing) %in% c("y"))]

#parameter choosing
error_testing<-c()
N<-c(300,400,500,600,700,800,900)

for(i in 1:7){

  fit <- train(training_x,training_y,N[i])

  pred_testing <- test(fit, testing_x)

  error_testing[i] <- mean(pred_testing != as.factor(testing_y))
  cat("Accurate rate for Training Random Forest with " ,N[i]," trees is ", 1-error_testing[i],'\n' )

}
```

Elapsed time for Training Random Forest with  300  trees is  28.4  seconds 
Accurate rate for Training Random Forest with  300  trees is  0.8544444 
Elapsed time for Training Random Forest with  400  trees is  37.22  seconds 
Accurate rate for Training Random Forest with  400  trees is  0.8533333 
Elapsed time for Training Random Forest with  500  trees is  47.33  seconds 
Accurate rate for Training Random Forest with  500  trees is  0.8522222 
Elapsed time for Training Random Forest with  600  trees is  54.82  seconds 
Accurate rate for Training Random Forest with  600  trees is  0.8555556 
Elapsed time for Training Random Forest with  700  trees is  66.46  seconds 
Accurate rate for Training Random Forest with  700  trees is  0.8533333 
Elapsed time for Training Random Forest with  800  trees is  73.55  seconds 
Accurate rate for Training Random Forest with  800  trees is  0.8511111 
Elapsed time for Training Random Forest with  900  trees is  84.57  seconds 
Accurate rate for Training Random Forest with  900  trees is  0.8544444 

Best parameter for HSV feature + RandomForest is 600.



```{r}
#rgb data
training <- read.csv("../output/rgb_training.csv")
testing <- read.csv("../output/rgb_testing.csv")

training_y <- training[ , "y"]
training_x <- training[ ,!(colnames(training) %in% c("y"))]
testing_y <- testing[ ,"y"]
testing_x <- testing[ , !(colnames(testing) %in% c("y"))]

#parameter choosing
error_testing<-c()
N<-c(300,400,500,600,700,800,900)

for(i in 1:7){

  fit <- train(training_x,training_y,N[i])

  pred_testing <- test(fit, testing_x)

  error_testing[i] <- mean(pred_testing != as.factor(testing_y))
  cat("Accurate rate for Training Random Forest with " ,N[i]," trees is ", 1-error_testing[i],'\n' )

}
```
Elapsed time for Training Random Forest with  300  trees is  115.56  seconds 
Accurate rate for Training Random Forest with  300  trees is  0.8677778 
Elapsed time for Training Random Forest with  400  trees is  158.03  seconds 
Accurate rate for Training Random Forest with  400  trees is  0.8677778 
Elapsed time for Training Random Forest with  500  trees is  193.66  seconds 
Accurate rate for Training Random Forest with  500  trees is  0.8755556 
Elapsed time for Training Random Forest with  600  trees is  232.16  seconds 
Accurate rate for Training Random Forest with  600  trees is  0.8666667 
Elapsed time for Training Random Forest with  700  trees is  274.35  seconds 
Accurate rate for Training Random Forest with  700  trees is  0.8733333 
Elapsed time for Training Random Forest with  800  trees is  305.82  seconds 
Accurate rate for Training Random Forest with  800  trees is  0.8711111 
Elapsed time for Training Random Forest with  900  trees is  357.05  seconds 
Accurate rate for Training Random Forest with  900  trees is  0.8744444 

Best parameter for RGB feature + RandomForest is 500.


```{r}
#sift data
library(dplyr)
sift_train <- read.csv("../data/training_set/SIFT_train.csv",header = F)
label_train<- read.csv("../data/training_set/label_train.csv")
training_y<- matrix(label_train[, 3],ncol=1)
colnames(training_y)<-"y"
training_x<- sift_train[, -1]
training1<- training_x
training1$y<- training_y
training<- sample_frac(training1, 0.7, replace=FALSE)
testing<- setdiff(training1, training, 'rows')

training_y <- training[ , "y"]
training_x <- training[ ,!(colnames(training) %in% c("y"))]
testing_y <- testing[ ,"y"]
testing_x <- testing[ , !(colnames(testing) %in% c("y"))]

#parameter choosing
error_testing<-c()
N<-c(300,400,500,600,700,800,900)

for(i in 1:7){

  fit <- train(training_x,training_y,N[i])

  pred_testing <- test(fit, testing_x)

  error_testing[i] <- mean(pred_testing != as.factor(testing_y))
  cat("Accurate rate for Training Random Forest with " ,N[i]," trees is ", 1-error_testing[i],'\n' )

}
```

Elapsed time for Training Random Forest with  300  trees is  258.63  seconds 
Accurate rate for Training Random Forest with  300  trees is  0.6688889 
Elapsed time for Training Random Forest with  400  trees is  347.91  seconds 
Accurate rate for Training Random Forest with  400  trees is  0.6777778 
Elapsed time for Training Random Forest with  500  trees is  430.23  seconds 
Accurate rate for Training Random Forest with  500  trees is  0.6688889 
Elapsed time for Training Random Forest with  600  trees is  517.32  seconds 
Accurate rate for Training Random Forest with  600  trees is  0.6711111 
Elapsed time for Training Random Forest with  700  trees is  606.16  seconds 
Accurate rate for Training Random Forest with  700  trees is  0.6711111 
Elapsed time for Training Random Forest with  800  trees is  688.15  seconds 
Accurate rate for Training Random Forest with  800  trees is  0.6722222 
Elapsed time for Training Random Forest with  900  trees is  775.59  seconds 
Accurate rate for Training Random Forest with  900  trees is  0.67 

Best parameter for SIFT feature + RandomForest is 400.


### Step3: Cross Validation Function 

```{r}

cv.function <- function(data, label, K,para){
  # data: the whole dataset
  # label: a column vector with 0 and 1
  # K: number of folds during the cross validation process

  set.seed(5243)
  library(caret)

  fold <- createFolds(1:nrow(data), K, list=T, returnTrain=F)
  fold <- as.data.frame(fold)
  
  cv.error <- c()
  
  for (i in 1:K){
    
    test.data <- data[fold[,i],]
    train.data <- data[-fold[,i],]
    test.label <- label[fold[,i],]
    train.label <- label[-fold[,i],]
    
    fit <- train(train.data, train.label,N=para)
    pred <- test(fit, test.data)  
    cv.error[i] <- mean(pred != test.label)
  }
  
   return(mean(cv.error))
}


```

### Feature selection

##### 1. full SIFT

```{r}
sift_train<- read.csv("../data/training_set/SIFT_train.csv",header = F)
label_train<- read.csv("../data/training_set/label_train.csv")
training_y<- matrix(label_train[, 3],ncol=1)
colnames(training_y)<-"y"
training_x<- sift_train[,-1]

cv.sift.rf<-cv.function(training_x,training_y,K=5,para = 400)

ar_cv.sift.rf<- 1-cv.sift.rf
ar_cv.sift.rf
```

Elapsed time for Training Random Forest with  400  trees is  464.16  seconds 
Elapsed time for Training Random Forest with  400  trees is  460.02  seconds 
Elapsed time for Training Random Forest with  400  trees is  467.97  seconds 
Elapsed time for Training Random Forest with  400  trees is  455.74  seconds 
Elapsed time for Training Random Forest with  400  trees is  458.23  seconds 

Accurate rate is 0.7036667.
Total time = 2306.12 second.

##### 2.HOG

```{r}
HOG_train <- rbind(read.csv("../output/hog_training.csv"),read.csv("../output/hog_testing.csv"))

training_y<- matrix(HOG_train[,ncol(HOG_train)],ncol=1)
colnames(training_y)<-"y"
training_x<- HOG_train [, -ncol(HOG_train)]

cv.HOG.rf<-cv.function(training_x,training_y,K=5,para = 500)

ar_cv.HOG.rf<-1-cv.HOG.rf

```
Elapsed time for Training Random Forest with  500  trees is  118.54  seconds 
Elapsed time for Training Random Forest with  500  trees is  120.28  seconds 
Elapsed time for Training Random Forest with  500  trees is  2220.36  seconds 
Elapsed time for Training Random Forest with  500  trees is  134.5  seconds 
Elapsed time for Training Random Forest with  500  trees is  123.4  seconds

Accurate rate is 0.8036667.
Total time = 2717.08 second.

```{r}
#rgb

RGB_train<- read.csv("../output/rgb_feature.csv")

training_y<- matrix(RGB_train[,ncol(RGB_train)],ncol=1)
colnames(training_y)<-"y"
training_x<- RGB_train[, -ncol(RGB_train)]

cv.RGB.rf<-cv.function(training_x,training_y,K=5,para = 500)

ar_cv.RGB.rf<-1-cv.RGB.rf


```
Elapsed time for Training Random Forest with  500  trees is  233.2  seconds 
Elapsed time for Training Random Forest with  500  trees is  230.94  seconds 
Elapsed time for Training Random Forest with  500  trees is  227.79  seconds 
Elapsed time for Training Random Forest with  500  trees is  227.76  seconds 
Elapsed time for Training Random Forest with  500  trees is  227.88  seconds 

Accurate rate is 0.8666667.
Total time = 1147.57 second

```{r}
lbp_train <- rbind(read.csv("../output/lbp_training.csv"),read.csv("../output/lbp_testing.csv"))
lbp_train <- lbp_train[,-1]
training_y<- matrix(lbp_train[,ncol(lbp_train)],ncol=1)
colnames(training_y)<-"y"
training_x<- lbp_train [, -ncol(lbp_train)]

cv.lbp.rf<-cv.function(training_x,training_y,K=5,para = 700)

ar_cv.lbp.rf<-1-cv.lbp.rf
ar_cv.lbp.rf
```
Elapsed time for Training Random Forest with  700  trees is  83.17  seconds 
Elapsed time for Training Random Forest with  700  trees is  84.95  seconds 
Elapsed time for Training Random Forest with  700  trees is  83.39  seconds 
Elapsed time for Training Random Forest with  700  trees is  83.58  seconds 
Elapsed time for Training Random Forest with  700  trees is  89.97  seconds 

Accurate rate is 0.8493333.
Total time = 425.06 second
